The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics

### 2016-02-01 Papers/InProgress/IJCAI/submission.tex
slight rewording in the abstract; added permission quote in Fig. 6

\begin{abstract}
  This paper discusses the discovery of the inconsistency in G\"odel's
  ontological argument as a success story for artificial
  intelligence. Despite the popularity of the argument since the
  appearance of G\"odel's manuscript in 1970, the inconsistency of the
  axioms used in the argument remained unnoticed until 2013, when it
  was detected automatically by the higher-order theorem prover
  \textsc{Leo-II}. 
  Understanding and verifying the refutation
  generated by the prover turned out to be a time-consuming task, and
  its completion, as reported here, required the reconstruction of the
  refutation in the Isabelle proof assistant and eventually led to a novel and more efficient way of automating higher-order modal logic \textbf{S5} with a universal accessibility relation. Furthermore, the development of an
  improved syntactical hiding for the embedding technique
  allows the refutation to be
  presented in a human-friendly way, suitable for non-experts in the
  technicalities of higher-order theorem proving. This brings us a
  step closer to wider adoption of logic-based artificial intelligence
  tools by philosophers.
\end{abstract}

@@

\subsection{Argument Reconstruction in Isabelle}  \label{sec:arg2}
\begin{figure}[t]
\centerline{\includegraphics[width=1\columnwidth]{./Images/InconsistencyIsabelleK.png}}
\caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
\end{figure}
\begin{figure*}[t]
  \centering
  \begin{subfigure}[t]{0.715\textwidth}
    \includegraphics[width=\textwidth]{./Images/Manuscript2.png}
    \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
      definitions highlighted} \label{GoedelScript} 
  \end{subfigure}
   \begin{subfigure}[t]{0.28\textwidth}
     \includegraphics[width=\textwidth,height=7.2cm]{./Images/Inconsistency_S5U_direct.png}
     \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
   \end{subfigure}
% \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
% \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
 \caption{The inconsistency in G\"{o}del's manuscript has been
   detected and verified by HOL ATPs} 
\end{figure*}
To verify the correctness of the informal argument explained above, it
was reconstructed in Isabelle/HOL, using Metis\footnote{Metis, unlike
  external provers such as \textsc{Leo-II} or Satallax, 
  constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
inessential parts (cf. Fig.~\ref {InconsistencyIsabelleK}). The essential use of the Empty Essence Lemma, on
the other hand, is explicitly stated, to ensure that Isabelle is
reconstructing the same argument. In fact, without the help of this
lemma, Metis is still not strong enough to refute G\"odel's
axioms.

>>>>

\begin{abstract}
  This paper discusses the discovery of the inconsistency in G\"odel's
  ontological argument as a success story for artificial
  intelligence. Despite the popularity of the argument since the
  appearance of G\"odel's manuscript in the early 70's, the
  inconsistency of the axioms used in the argument remained unnoticed
  until 2013, when it was detected automatically by the higher-order
  theorem prover \textsc{Leo-II}.  Understanding and verifying the
  refutation generated by the prover turned out to be a time-consuming
  task. Its completion, as reported here, required the
  reconstruction of the refutation in the Isabelle proof assistant, and it
  also led to a novel and more efficient way of automating
  higher-order modal logic \textbf{S5} with a universal accessibility
  relation. Furthermore, the development of an improved syntactical
  hiding for the utilized logic embedding technique allows the refutation to be
  presented in a human-friendly way, suitable for non-experts in the
  technicalities of higher-order theorem proving. This brings us a
  step closer to wider adoption of logic-based artificial intelligence
  tools by philosophers.
\end{abstract}

@@

\subsection{Argument Reconstruction in Isabelle}  \label{sec:arg2}
\begin{figure}[t]
\centerline{\includegraphics[width=1\columnwidth]{./Images/InconsistencyIsabelleK.png}}
\caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
\end{figure}
\begin{figure*}[t]
  \centering
  \begin{subfigure}[t]{0.715\textwidth}
    \includegraphics[width=\textwidth]{./Images/Manuscript2.png}
    \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
      definitions highlighted (with permission from the Kurt G\"odel Papers, Shelby White and Leon Levy Archives Center, Princeton, NJ, USA, on deposit at Princeton University)} \label{GoedelScript} 
  \end{subfigure}
   \begin{subfigure}[t]{0.28\textwidth}
     \includegraphics[width=\textwidth,height=7.2cm]{./Images/Inconsistency_S5U_direct.png}
     \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
   \end{subfigure}
% \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
% \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
 \caption{The inconsistency in G\"{o}del's manuscript has been
   detected and verified by HOL ATPs} 
\end{figure*}
To verify the correctness of the informal argument explained above, it
was reconstructed in Isabelle/HOL, using Metis\footnote{Metis, unlike
  external provers such as \textsc{Leo-II} or Satallax, 
  constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
inessential parts (cf. Fig.~\ref {InconsistencyIsabelleK}). The essential use of the Empty Essence Lemma, on
the other hand, is explicitly stated, to ensure that Isabelle is
reconstructing the same argument. In fact, without the help of this
lemma, Metis is still not strong enough to refute G\"odel's
axioms.
